Place of death in NRS COVID-19 deaths: data processing for analysis

### 2022-01-13 covid_nrs_place_of_death_data.Rmd
tidied some commented out stuff

# TODO:  still too complicated to safely use as a function...
## find which combinations of geographic area, cause of death, and place of death are missing and set the number of deaths to 0 for them
# complete_weekly_deaths_with_implicit_zeros <- function(data_tbl) {
#   data_tbl %>%
#   tidyr::complete(
#       nesting(year, week_number, week_number_run_over, date_w_c),  # these are 'identifying' variables that shouldn't be combinatorily exploded
#       ref_area, cause_of_death, place_of_death,  # these are the variables we want to have all combinations of, and missing combinations are set to 0 deaths
#       fill = list(number_of_deaths = 0)
#       )
# }

@@

## covid death reporting started in week 12, 2020 - before then it doesn't make sense to distunguish between covid and non-covid deaths, and only the all cause death figure should be kept
## by using the 'overrun' weeks we ensure that 2021 or later isn't affected
## note: we're keeping all cause deaths!
set_covid_and_non_covid_deaths_to_zero_before_reporting_started <- function(data_tbl) {
  data_tbl %>%
    mutate(
      deaths_non_covid = if_else(
        condition = week_number_run_over < 12,
        true = NA_real_,
        false = deaths_non_covid
      ),
      deaths_covid_related = if_else(
        condition = week_number_run_over < 12,
        true = NA_real_,
        false = deaths_covid_related
      )
    )
}

# TODO: delete below function, it's too complicated and I've broken it down to the individual functions
# recode_cause_of_death_and_compute_columns <- function(data_tbl) {
#   potential_grouping_variables <- c("ref_area","sex","age","ref_period")
#   grouping_variables <- potential_grouping_variables[which(potential_grouping_variables %in% names(data_tbl))]
#   
#   interim_data_tbl <- 
#     data_tbl %>%
#     recode_cause_of_death() %>%
#     tidyr::complete(
#       nesting(year, week_number, week_number_run_over, date_w_c),
#       !!!rlang::syms(grouping_variables),
#       cause_of_death, place_of_death,
#       fill = list(number_of_deaths = 0)
#       ) %>%
#     pivot_wider(names_from = "cause_of_death", values_from = "number_of_deaths")
#   
#   ## compute non-covid deaths from all deaths minus covid deaths
#   if (!"deaths_non_covid" %in% names(interim_data_tbl)) {
#     interim_data_tbl <-
#       interim_data_tbl %>%
#       mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related)
#   }
#   ## compute all deaths from non-covid deaths + covid deaths
#   if (!"deaths_all_causes" %in% names(interim_data_tbl)) {
#     interim_data_tbl <-
#       interim_data_tbl %>%
#       mutate(deaths_all_causes = deaths_non_covid + deaths_covid_related)
#   }
#   
#   interim_data_tbl <-
#     interim_data_tbl %>%  # set as missing covid- & non-covid-related deaths before week 12
#     mutate(
#       deaths_non_covid = if_else(
#         condition = week_number_run_over < 12,
#         true = NA_real_,
#         false = deaths_non_covid
#       ),
#       deaths_covid_related = if_else(
#         condition = week_number_run_over < 12,
#         true = NA_real_,
#         false = deaths_covid_related
#       )
#     )
#   
#   return(interim_data_tbl)
# }
```

>>>>

@@

## covid death reporting started in week 12, 2020 - before then it doesn't make sense to distunguish between covid and non-covid deaths, and only the all cause death figure should be kept
## by using the 'overrun' weeks we ensure that 2021 or later isn't affected
## note: we're keeping all cause deaths!
set_covid_and_non_covid_deaths_to_zero_before_reporting_started <- function(data_tbl) {
  data_tbl %>%
    mutate(
      deaths_non_covid = if_else(
        condition = week_number_run_over < 12,
        true = NA_real_,
        false = deaths_non_covid
      ),
      deaths_covid_related = if_else(
        condition = week_number_run_over < 12,
        true = NA_real_,
        false = deaths_covid_related
      )
    )
}
```
